Design and Implementation of Probabilistic Programming Language Anglican

### 2017-02-27 paper/paper.tex
last changes forgot to commit

\usepackage{siunitx}
\usepackage[color]{changebar}
\usepackage{enumitem}
\usepackage[lighttt]{lmodern}
\usepackage{graphicx}

\usepackage{listings}

\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}
\lstdefinestyle{default}{language=Anglican,basicstyle=\ttfamily, columns=flexible, showstringspaces=false}

\usepackage{url}
\usepackage{etoolbox}

@@

Having defined the query, we are now ready to run the query
using an inference algorithm. Function
\texttt{doquery}\iftoggle{full}{ provided
by the \texttt{anglican.core} namespace}{} accepts the inference
algorithm, the query, and optional parameters, and returns a
lazy sequence of samples. We use here the inference algorithm
called Lightweight Metropolis-Hastings (LMH). The algorithm is
somewhat slow to converge but can be used with any Anglican
query, and should be robust enough for our simple problem.

We bind the results of \texttt{doquery} to variable
\texttt{samples}, to analyse the results later. However, since
the sequence is lazy, no inference is performed and no samples
are generated until they are retrieved and processed.

@@

To approximate the inferred distribution, we extract a finite
subsequence of samples. Many algorithms use an initial
subsequence of samples to converge to the target distribution.
Hence, we drop initial $N$ samples ($N$ is 5000 in the code
snippet below), and collect the next $N$ samples.

@@

Besides computing the posterior probability that the same
customer visited the deli twice, we may want to know the average
time (or times) to arrive. In Bayesian inference, it is common
to report distributions instead of `most likely' values. We use
query results in retrieved samples to approximate the
distributions, and plot distribution histograms for the same
customer visiting twice (Figure~\ref{fig:dist-same}) and
two different customers (Figure~\ref{fig:dist-diff})

\begin{lstlisting}[style=default]
;; single customer                      
(def time-to-arrive+ 
    (map (fn [x] (first (:times-to-arrive x)))
         (filter :same-customer results)))
(def mean-to-arrive+ (mean time-to-arrive+))
(def sd-to-arrive+ (std time-to-arrive+))

;; two customers
(def times-to-arrive+ 
     (map :times-to-arrive 
          (filter  
            (fn [x] (not (:same-customer x)))
            results)))
(def mean-1-to-arrive+
     (mean (map first times-to-arrive+)))
(def sd-1-to-arrive+
     (std (map first times-to-arrive+)))
(def mean-2-to-arrive+
     (mean (map second times-to-arrive+)))
(def sd-2-to-arrive+
     (std (map second times-to-arrive+)))
\end{lstlisting}

In addition to plotting the distribution histograms, we use
functions \texttt{mean}, \texttt{std} provided
in the \texttt{anglican.{\linebreak[0]}stat} namespace
along with other useful statistical functions to compute
summary statistics. This is another illustration of advantage of
tight integration between Clojure and Anglican --- probabilistic
models are expressed in Anglican. However, processing of data
and results can rely on the full power of Clojure.

>>>>

\usepackage{siunitx}
\usepackage[color]{changebar}
\usepackage{enumitem}
\usepackage[lighttt]{lmodern}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{url}
\usepackage{etoolbox}

\ttfamily
\DeclareFontShape{OT1}{lmtt}{m}{it}
     {<->sub*lmtt/m/sl}{}
\iftoggle{full}{
\lstdefinestyle{default}{language=Anglican,basicstyle=\ttfamily\small, columns=flexible, showstringspaces=false}
}{
\lstdefinestyle{default}{language=Anglican,basicstyle=\linespread{0.85}\ttfamily\small, columns=flexible, showstringspaces=false}
}

@@

Having defined the query, we are now ready to run the query
using an inference algorithm. Function
\texttt{doquery}\iftoggle{full}{ provided
by the \texttt{anglican.core} namespace}{} accepts the inference
algorithm, the query, and optional parameters, and returns a
lazy sequence of samples. We use here the inference algorithm
called Lightweight Metropolis-Hastings (LMH). \iftoggle{full}{The algorithm is
somewhat slow to converge but can be used with any Anglican
query, and should be robust enough for our simple problem.}{}
We bind the results of \texttt{doquery} to variable
\texttt{samples}, to analyse the results later. However, since
the sequence is lazy, no inference is performed and no samples
are generated until they are retrieved and processed.

@@

To approximate the inferred distribution, we extract a finite
subsequence of samples. \iftoggle{full}{Many algorithms use an initial
subsequence of samples to converge to the target distribution.
Hence, we drop initial $N$ samples ($N$ is 5000 in the code
snippet below), and collect the next $N$ samples.}{}

@@

Besides computing the posterior probability that the same
customer visited the deli twice, we may want to know the average
time (or times) to arrive. In Bayesian inference, it is common
to report distributions instead of `most likely' values. We use
query results in retrieved samples to approximate the
distributions, and plot distribution histograms for the same
customer visiting twice (Figure~\ref{fig:dist-same}) and
two different customers (Figure~\ref{fig:dist-diff}).

\iftoggle{full}{%
\input{case-study-mean-std}
}{}%
